%!TEX root = ../main.tex
%!TEX root = ../main.tex
\section{Mining Background}
Mining is a key part of this thesis, while not an in depth knowledge of mining is required, basic information is needed for problem definitions and clarity of comminication. The focus of the thesis is the removal of overburden from strip mines using draglines, as this is a specific topic and not a general discussion on optimisation of mining machines in general only infomration on draglines and the enviroment in which they operate should be considered. 
\subsection{Strip Mining}
\subsection{Draglines}
\subsection{Blocks}

\section{Operations Research Background}
The field of operations research is a mathematical discipline specailising in the application of advanced analytical methods to make better decisions for specified problems. Often considered as a subsection of applied mathmatics operations research employs techniques from mathmatical modelling and optimization theory to calculate optimal or near optimal solutions to complex decision making problems. \\ Naturally this branch of mathematics lends itself to the problem of optimal block dimensions in a strip mine, with the use of optimal decisions through either precise calculation or policy the best possible solution can be found using the field of operations research. The scope of this field is diverse, with many potential solutions to a problem, however the generation of the matematical model will aid in selecting a method that will obtain results quickest. 
\\
Mathematical modelling is a vital section of operations research, however the problem can be written and formulated in many different ways, as such the specific formulation styles of a technique will also be mentioned in that section, modelling in general can follow some simple rules yet no overall method or technique exists for the creation of a mathematical model, it is instead reliant on understanding of the system. 

\subsection{Linear Programming}
Linear programming is a technique used to obtain the optimal outcome of a model such as the minimal time or highest profit. This method can only be aplied to a mathematical model with a linear objective function subject to linear equality and linear inequalitiy constrants.\\ Often the model is described as groups of variables and contstraints with an overall objective function. This method of optimisation is often solved using complex optimisation engines such as gurobi rather than through algorithms implemented specifically for the implementation. The process of optimisation is done by defining a feasible region of valid solutions as a convex $n^{th}$ dimensional polytope, where $n$ is the dimensionality of the variables associated. Here this polytope is defined as an intersection of a finite amount of half spaces defined by linear inequalities. Within this polytope the optimal feasible solution will be found on a vertex satisfying all constraints and the objective function, this method can be calculated using matrix manipulation in algorithms such as the modified simplex algorithm.  	
\\
The standard variables ina linear program are continous, however a subtechniqe of linear programming is integer programming which adds the constraint that some variables must be full positive integers. This problem is typically harder to solve as the optimal linear solution may not be similar to the optimal integer solution. An integer solution is often calculated by calculating an optimal linear solution and branching on rounding decisions for the integer variables to create a decision tree. This tree is branched such that the difference between the linear solution and chosen integer solution is minimised, this is the origin of the increased difficulty of integer programming in comparison to linear programming. 

\subsection{Dynamic Programming} 
Dynamic programming is a method of optimisation in which a subproblem is solved and the solution stored, this is done for all subproblems until the full problem can be constructed from, if a subproblem has already been solved and stored then it is called from memory instead of through calculation. The process of storing sub problems for later use is called memotization and is vital for use in dynamic programming as dynammic programming often encounters similar subproblems when working recursivly. \\
typically a dynamic program refers to recursivly breaking down decisoions into steps through a stage. The value of each decision in a stage is a combination of the intial myopic value of the choice as well as the value of all other solutions that occur after this decision is made. This requires that some state in the problem is altered when a decision is made, then the affects of this decision must also be considered for all subsequent choices. This creates and expanding tree of subproblems and is why the method of memotisation is so vital to dynamic programming as often the same subproblem will be encountered multiple times. 
\subsection{Metaheuristics} %Should I mention this if I looked into it but never used it
A metahueristic is a higher level opimisation technique used to find, generate or select a hueristic that may provide a good solution to an optimisation problem, typically when working with incomplete information , unaccurate information or limited computational power. Unlike Linear programming and dynamic progamming a metahueristic does not garuntee a globally optimal solution , due to metahueristics typically relying on stochastic optimisation to generate solutions. \\
The genetic algorithm is such an example of a metahueristic where canddate solutions are evolved to find a better solution, this is done by calculating the value of all candidates. After this is complete the more preffereable solutions are combined with one another in such a way that a new series of candidate solutions is generated. This is done iteraitvly until the change between generations is suffciently small, at this point either a local optimal or global optimal solution can be found. Typically the genetic algorithm is used for determining a policy for good solutions, rather than linear programming or dynamic programming which will calculate the true optimal solution. 
